Developmental trajectories of bilingual word recognition

### 2023-10-13 manuscript/manuscript.tex

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, frame hidden, sharp corners, enhanced, interior hidden, breakable]}{\end{tcolorbox}}\fi

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The formation of a mental lexicon is a critical developmental
achievement for infants\_ learning words allows infants to retrieve
socially relevant concepts from fairly arbitrary linguistic forms
embedded in speech.

Learning words allows infants to access the rich world made of concepts
through the recognition of fairly arbitrary linguistic forms, i.e.,
words. This remarkable developmental achievement is

The bilingual lexicon is language-non selective.

The role of cognates in lexical processing.

How does language co-activation shape lexical development?

Implicit naming as a paradigm to study cross-language activation. (Mani
\& Plunkett, 2010, 2011a).

@@

All materials, data, and reproducible code can be found at the OSF
(\href{https://osf.io/ckydb/}{https://osf.io/hy984/}) and GitHub
(\url{https://github.com/gongcastro/cognate-priming}) repositories. This
study was conducted according to guidelines laid down in the Declaration
of Helsinki, and was approved by the Drug Research Ethical Committee
(CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before
every testing session, caretakers were asked to read and sign an
informed consent form, and were given a small gift at the end of it

@@

\NormalTok{n\_sessions\_age\_group }\OtherTok{\textless{}{-}}\NormalTok{ participants }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_by}\NormalTok{(age\_group) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(age, }\FunctionTok{lst}\NormalTok{(mean, sd, min, max)),}
              \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
              \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(age\_mean}\SpecialCharTok{:}\NormalTok{age\_max, round, }\DecValTok{2}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_split}\NormalTok{(age\_group) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{set\_names}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"age\_21"}\NormalTok{, }\StringTok{"age\_25"}\NormalTok{, }\StringTok{"age\_30"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{verbatim}
Warning: [1m[22mThere was 1 warning in `mutate()`.
[1m[22m[36mi[39m In argument: `across(age_mean:age_max, round, 2)`.
Caused by warning:
[1m[22m[33m![39m The `...` argument of `across()` is deprecated as of dplyr 1.1.0.
Supply arguments directly to `.fns` through an anonymous function instead.

  # Previously
  across(a:b, mean, na.rm = TRUE)

  # Now
  across(a:b, \(x) mean(x, na.rm = TRUE))
\end{verbatim}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n\_sessions\_dominance }\OtherTok{\textless{}{-}} \FunctionTok{count}\NormalTok{(participants, test\_language) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_split}\NormalTok{(test\_language) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{set\_names}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"catalan"}\NormalTok{, }\StringTok{"spanish"}\NormalTok{))}

@@

We collected vocabulary data using parental responses to the Barcelona
Vocabulary Inventory (BVQ,
\protect\hyperlink{ref-garcia-castro2023bvq}{\textbf{garcia-castro2023bvq?}}),
an online vocabulary checklist inspired in several adaptations of the
the Communicative Developmental Inventory (CDI,
\protect\hyperlink{ref-fenson2004variability}{\textbf{fenson2004variability?}})
developed to assess the vocabulary size of Catalan-Spanish bilingual
toddlers. Families received a link to the BVQ immediately after each
experimental session, and were given two weeks to fill it. Two weeks
after being sent the link, 141 (48\% families failed to provide a
complete response to the BVQ. We imputed their vocabulary size scores
using multiple imputation
(\protect\hyperlink{ref-vanbuuren2018flexible}{\textbf{vanbuuren2018flexible?}}).
We used predictive mean matching to impute missing scores using
participants' information about their age group and language profile. We
obtained several receptive and expressive vocabulary size measures from
each participant's vocabulary: Catalan vocabulary size (proportion of
the words in the Catalan checklist reported as acquired), Spanish
vocabulary size (proportion of the words in the Catalan checklist
reported as acquired), and total vocabulary size (proportion of the
words in both checklists reported as acquired).

\hypertarget{stimuli}{%
\subsubsection{Stimuli}\label{stimuli}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n\_words }\OtherTok{\textless{}{-}} \FunctionTok{unique}\NormalTok{(}\FunctionTok{with}\NormalTok{(stimuli, }\FunctionTok{unlist}\NormalTok{(prime, target, distractor)))}
\end{Highlighting}
\end{Shaded}

We used arbre, ànec, camió, finestra, vaca, barret, osset, claus, cotxe,
cadira, vaixell, ampolla, flor, formatge, botó, ulleres, gat, gos, mico,
pa, moto, mitjó, avió, granota, ovella, ocell, peix, plàtan, pilota,
poma, guitarra, maduixa, botella, globo, mono, coche, cama, zumo, pato,
cebra, cerdo, oreja, mesa, galleta, gafas, tomate, zapato, manzana,
vaso, queso, oveja, ojo, pelota, tele, pez, perro, barco, hoja, tren,
tenedor, casa, gato distinct words included in the BVQ to create the
stimuli lists. We created six stimuli lists: three in Catalan, and three
in Spanish. Each list contained 32 trials, each involving a
prime-target-distractor group. Each word played a role as either prime,
\emph{or} as target and distractor across the three lists in their
corresponding language. For instance, the Catalan word \emph{cadira}
appeared as \emph{prime} in the three lists, but never as \emph{target}
or \emph{distractor}; the Catalan word \emph{bici} appeared as
\emph{target} and \emph{distractor} across the three lists, but never as
a prime. Target-distractor pairings were held constant across the three
lists in each language. For instance, in all Catalan lists the word
\emph{bici} was paired with the word \emph{porta}. Target-distractor
pairings were also yoked, so that each member of the same
target-distractor pair appeared once as target and once as distractor in
each list. For instance, the \emph{bici}-\emph{porta} paired appeared
twice in each of the three Catalan lists: once with \emph{bici} as
target and \emph{porta} as distractor, and once with \emph{porta} as
target and \emph{bici} as distractor. This counterbalancing avoided
participants encountering looking at the target word guided solely by
that word having being named in a previous trial. Finally, prime words
appeared only once in each list: each target-distractor pair was
associated with a different prime word in both appearances. In each
list, the same prime word was presented alongside a different
target-distractor pair. For instance, the Catalan prime word
\emph{barret} was presented with the \emph{bici}-\emph{porta}
target-distractor pair in one list, with the \emph{bici}-\emph{porta}
pair in another list, and with \emph{berenar}-\emph{amanida} in the
remaining list. The order of the trials was randomised across
experimental session, so that each time a participant was tested, the
order in which the prime-target-distractor was presented was randomised.
Each participant was randomly assigned to one of the three list in the
corresponding language (their dominant language, see
Section~\ref{sec-lp}), and always the same list across their
experimental sessions in the case of a recurrent participant.

In 16 of the 32 trials of the same list (henceforth \emph{related}
trials), the prime and the target words were phonologically related,
sharing phonological onset (at least first phoneme). In the other 16
trials (\emph{unrelated} trials), prime and target did not share
phonological onset. 8 of the 16 \emph{related} trials included a cognate
prime (\emph{cognate} trials), and the other 8 included a non-cognate
trials (\emph{non-cognate} trials). A prime word was considered cognate
if its Catalan and Spanish translation shared phonological onset.
Especial attention was paid to avoiding semantic or taxonomic
relationships between prime and target words, and between prime and
distractor words. Target and distractor word pairs were phonologically
unrelated (did not share phonological onset). Some of them shared
semantic features or a taxonomic relationship. This is the case of words
associated with especially salient referents such as animals or food. To
avoid infants guiding their gaze to these objects based on their
saliency, we paired animals and food items together.

@@

\hypertarget{auditory-stimuli}{%
\paragraph{Auditory stimuli}\label{auditory-stimuli}}

The auditory stimuli were natural exemplars of the selected target
words, spoken by Catalan-Spanish proficient bilingual female speaker who
was instructed to pronounce each word in a toddler-directed manner.
Recordings were made with an Audio-Tecnica 328 microphone (AT2050) at a
sampling rate of 44100 Hz, in a soundproof room at the \emph{Laboratori
de Recerca en Infancia} at University Pompeu Fabra. We used the Audacity
(\protect\hyperlink{ref-reference}{\textbf{reference?}}) and Praat
(\protect\hyperlink{ref-reference}{\textbf{reference?}}) to record and
edit the audio files. The speaker was presented with a list of words in
Catalan. The order of the words was pseudo-randomised, and each word was
produced three times in a row before moving to the next word in the
list. After going through all the words in the list, the speaker went
through the word list again generating three tokens for each word, now
in an inverse order (from bottom of the list to the top). We then
repeated the same procedure for the list of Spanish words. The resulting
audios were manually chunked into individual word-forms. For each of the
six tokens produced for each word, the most adequate was selected for
further processing. The audios were then transformed to stereo by
duplicating them into two channels, denoised, and finally normalised.
The mean duration of the final audios was 1.23 (\emph{SD} = 0.17) and
1.08 (\emph{SD} = 0.14) seconds for the Catalan and Spanish lists.

@@

\hypertarget{visual-stimuli}{%
\paragraph{Visual stimuli}\label{visual-stimuli}}

For each word, we created a picture with a typical referent.

@@

\hypertarget{data-processing}{%
\subsubsection{Data processing}\label{data-processing}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n\_trials }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(attrition\_trials)}

\NormalTok{n\_sessions }\OtherTok{\textless{}{-}} \FunctionTok{distinct}\NormalTok{(attrition\_trials, filename) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{nrow}\NormalTok{()}

\NormalTok{n\_participants }\OtherTok{\textless{}{-}} \FunctionTok{distinct}\NormalTok{(attrition\_trials, id) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{nrow}\NormalTok{()}

\NormalTok{n\_trials\_prime }\OtherTok{\textless{}{-}}\NormalTok{ attrition\_trials }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{filter}\NormalTok{(is\_valid\_gaze\_prime)}

\NormalTok{n\_trials\_test }\OtherTok{\textless{}{-}}\NormalTok{ n\_trials\_prime }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{filter}\NormalTok{(is\_valid\_gaze\_test)}

\NormalTok{n\_trials\_each }\OtherTok{\textless{}{-}}\NormalTok{ n\_trials\_test }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{filter}\NormalTok{(is\_valid\_gaze\_test\_each)}

\NormalTok{n\_trials\_vocab }\OtherTok{\textless{}{-}}\NormalTok{ n\_trials\_each }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{filter}\NormalTok{(is\_valid\_vocab)}

\NormalTok{n\_trials\_participant }\OtherTok{\textless{}{-}}\NormalTok{ n\_trials\_vocab }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{inner\_join}\NormalTok{(}\FunctionTok{filter}\NormalTok{(attrition\_participants,}
\NormalTok{                      is\_valid\_participant),}
               \AttributeTok{by =} \FunctionTok{join\_by}\NormalTok{(filename))}

\NormalTok{n\_sessions\_valid }\OtherTok{\textless{}{-}}\NormalTok{ attrition\_participants }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{filter}\NormalTok{(is\_valid\_participant)}

\NormalTok{n\_participants\_valid }\OtherTok{\textless{}{-}}\NormalTok{ n\_trials\_participant }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{distinct}\NormalTok{(filename)}

\NormalTok{n\_longitudinal }\OtherTok{\textless{}{-}} \FunctionTok{count}\NormalTok{(n\_sessions\_valid, filename, }\AttributeTok{name =} \StringTok{"sessions"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{count}\NormalTok{(sessions)}

\NormalTok{n\_participants\_include }\OtherTok{\textless{}{-}} \FunctionTok{read\_csv}\NormalTok{(}\StringTok{"data{-}raw/participants.csv"}\NormalTok{,}
                                   \AttributeTok{show\_col\_types =} \ConstantTok{FALSE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

We gathered data for 9,472 trials from 296 experimental sessions,
provided by 180 distinct participants. Missing eye-tracker samples were
interpolated using the last-observation-carried-forward (see
\protect\hyperlink{ref-zettersten2022peekbank}{\textbf{zettersten2022peekbank?}}
for a similar approach). We excluded trials in which the participant did
not fixate the prime at least half of the time of the prime phase
(\$\geq\(750 ms, *n* = 1,810), or the target and distractor during the test phase (\)\geq\$1,000
ms, \emph{n} = 502). We also excluded trials in which the participant
failed to fixate both target \emph{and} distractor pictures for at least
100 ms (i.e., 5\% of the time) (1,304 trials excluded). We then excluded
trials in which the participant was not reported to understand the prime
\emph{and} the target words, according to their caregivers' responses to
the vocabulary checklist. This resulted in the exclusion of 3,641
trials. Finally, we excluded trials from experimental sessions in which
the participant did not provide at least two valid trials in each of the
three experimental conditions (cognate prime, non-cognate prime, and
unrelated prime) which resulted in the exclusion of 34 experimental
sessions and their 219 remaining trials. The final data set comprised 0
from c(``CognatePriming\_019\_2019-11-28.csv'',
``CognatePriming\_021\_2019-11-30.csv'',
``CognatePriming\_026\_2019-12-14.csv'',
``CognatePriming\_027\_2019-12-14\_1.csv'',
``CognatePriming\_049\_2021-02-13.csv'',
``CognatePriming\_050\_2020-11-07.csv'',
``CognatePriming\_050\_2021-03-13.csv'',
``CognatePriming\_050\_2021-10-02.csv'',
``CognatePriming\_051\_2021-03-06\_1.csv'',
``CognatePriming\_051\_2021-10-01.csv'',
``CognatePriming\_053\_2021-04-17.csv'',
``CognatePriming\_054\_2020-11-14.csv'',
``CognatePriming\_055\_2020-11-14.csv'',
``CognatePriming\_055\_2021-02-20\_1.csv'',
``CognatePriming\_056\_2020-11-14.csv'',
``CognatePriming\_058\_2020-11-17.csv'',
``CognatePriming\_059\_2020-11-18.csv'',
``CognatePriming\_060\_2021-04-26.csv'',
``CognatePriming\_064\_2020-11-21.csv'',
``CognatePriming\_065\_2020-11-25.csv'',
``CognatePriming\_066\_2020-11-26.csv'',
``CognatePriming\_067\_2020-11-27.csv'',
``CognatePriming\_068\_2020-11-27.csv'',
``CognatePriming\_069\_2020-11-28.csv'',
``CognatePriming\_069\_2021-05-08.csv'',
``CognatePriming\_070\_2020-12-01.csv'',
``CognatePriming\_070\_2021-09-30\_1.csv'',
``CognatePriming\_071\_2020-12-01.csv'',
``CognatePriming\_072\_2020-12-04.csv'',
``CognatePriming\_073\_2021-10-02.csv'',
``CognatePriming\_074\_2020-12-05.csv'',
``CognatePriming\_075\_2020-12-09.csv'',
``CognatePriming\_076\_2021-05-08.csv'',
``CognatePriming\_076\_2021-10-16.csv'',
``CognatePriming\_078\_2021-04-29.csv'',
``CognatePriming\_079\_2020-12-16.csv'',

>>>>

\begin{document}
\maketitle
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[interior hidden, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}, breakable, boxrule=0pt, enhanced]}{\end{tcolorbox}}\fi

@@

All materials, data, and reproducible code can be found at the OSF
(\href{https://osf.io/ckydb/}{https://osf.io/hy984/}) and GitHub
(\url{https://github.com/gongcastro/cognate-priming}) repositories. This
study was conducted according to guidelines laid down in the Declaration
of Helsinki, and was approved by the Drug Research Ethical Committee
(CEIm) of the IMIM Parc de Salut Mar, reference 2020/9080/I. Before
every testing session, caregivers were asked to read and sign an
informed consent form, and were given a token of appreciation at the end
of it.

@@

\NormalTok{n\_sessions\_age\_group }\OtherTok{\textless{}{-}}\NormalTok{ participants }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_by}\NormalTok{(age\_group) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(age, }\FunctionTok{lst}\NormalTok{(mean, sd, min, max)),}
              \AttributeTok{n =} \FunctionTok{n}\NormalTok{(),}
              \AttributeTok{.groups =} \StringTok{"drop"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{mutate}\NormalTok{(}\FunctionTok{across}\NormalTok{(age\_mean}\SpecialCharTok{:}\NormalTok{age\_max, \textbackslash{}(x) }\FunctionTok{round}\NormalTok{(x, }\DecValTok{2}\NormalTok{))) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_split}\NormalTok{(age\_group) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{set\_names}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"age\_21"}\NormalTok{, }\StringTok{"age\_25"}\NormalTok{, }\StringTok{"age\_30"}\NormalTok{))}

\NormalTok{n\_sessions\_dominance }\OtherTok{\textless{}{-}} \FunctionTok{count}\NormalTok{(participants, test\_language) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{group\_split}\NormalTok{(test\_language) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{set\_names}\NormalTok{(}\FunctionTok{c}\NormalTok{(}\StringTok{"catalan"}\NormalTok{, }\StringTok{"spanish"}\NormalTok{))}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n\_imputed }\OtherTok{\textless{}{-}} \FunctionTok{table}\NormalTok{(vocabulary}\SpecialCharTok{$}\NormalTok{is\_imputed)[}\DecValTok{2}\NormalTok{]}
\NormalTok{prop\_imputed }\OtherTok{\textless{}{-}}\NormalTok{ scales}\SpecialCharTok{::}\FunctionTok{percent}\NormalTok{(n\_imputed}\SpecialCharTok{/}\FunctionTok{nrow}\NormalTok{(vocabulary))}
\NormalTok{n\_pool }\OtherTok{\textless{}{-}} \FunctionTok{nrow}\NormalTok{(bvq\_data}\SpecialCharTok{$}\NormalTok{vocabulary)}
\end{Highlighting}
\end{Shaded}

We collected vocabulary data using parental responses to the Barcelona
Vocabulary Inventory (BVQ,
\protect\hyperlink{ref-garcia-castro2023bvq}{Garcia-Castro et al.,
2023}), an online vocabulary checklist inspired in several adaptations
of the the Communicative Developmental Inventory (CDI,
\protect\hyperlink{ref-fenson1994variability}{Fenson et al., 1994})
developed to assess the vocabulary size of Catalan-Spanish bilingual
toddlers. Families received a link to the BVQ immediately after each
experimental session, and were given two weeks to fill it.

We calculated several measures of receptive vocabulary size from each
participant's vocabulary: L1 vocabulary size (proportion of words
reported as acquired in the checklist of the dominant language), L2
vocabulary size (proportion of words reported as acquired in the
checklist of the non-dominant language), total vocabulary size
(proportion of the words in both checklists reported as acquired),
conceptual vocabulary (proportion of concepts for which the participant
was reported to have acquired at least one label, in any language), and
translation equivalent vocabulary (proportion of concepts for which the
participant was reported to have acquired at two labels, one in each
language).

141 (48\%) Families failed to provide a complete response to the BVQ
within the two-week time limit, or did not provide a successful response
to the questionnaire. For missing questionnaire responses, we imputed
the vocabulary size of the participant using single imputation, using
the vocabulary size scores of a pool of 600 additional participants for
which a successful response for the questionnaire had been gathered. We
used participants age in months and their language profile
(monolingual/bilingual) as predictors. We used the \texttt{mice} R
package (\protect\hyperlink{ref-van2011mice}{Van Buuren \&
Groothuis-Oudshoorn, 2011}) to perform imputation using the Bayesian
linear regression method.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{vocabulary }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{left\_join}\NormalTok{(}\FunctionTok{select}\NormalTok{(participants, filename, id, age\_group, lp),}
              \AttributeTok{by =} \FunctionTok{join\_by}\NormalTok{(filename)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{summarise}\NormalTok{(}\FunctionTok{across}\NormalTok{(}\FunctionTok{matches}\NormalTok{(}\StringTok{"prop"}\NormalTok{), }
\NormalTok{                     tibble}\SpecialCharTok{::}\FunctionTok{lst}\NormalTok{(mean, sd)),}
              \AttributeTok{.by =} \FunctionTok{c}\NormalTok{(age\_group, lp)) }\SpecialCharTok{|\textgreater{}}
    \FunctionTok{arrange}\NormalTok{(age\_group, lp) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{gt}\NormalTok{(}\AttributeTok{groupname\_col =} \StringTok{"age\_group"}\NormalTok{,}
       \AttributeTok{rowname\_col =} \StringTok{"lp"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_hide}\NormalTok{(lp) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"Total "}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"total"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"L1"}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"l1"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"L2"}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"l2"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"Conceptual"}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"concept"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"TE"}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"te\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_spanner}\NormalTok{(}\StringTok{"Vocabulary size (\%)"}\NormalTok{, }\FunctionTok{matches}\NormalTok{(}\StringTok{"prop\_"}\NormalTok{)) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_merge\_uncert}\NormalTok{(total\_prop\_mean, total\_prop\_sd) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_merge\_uncert}\NormalTok{(l1\_prop\_mean, l1\_prop\_sd) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_merge\_uncert}\NormalTok{(l2\_prop\_mean, l2\_prop\_sd) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_merge\_uncert}\NormalTok{(concept\_prop\_mean, concept\_prop\_sd) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_merge\_uncert}\NormalTok{(te\_prop\_mean, te\_prop\_sd) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{fmt\_number}\NormalTok{(is.numeric, }
               \AttributeTok{drop\_trailing\_zeros =} \ConstantTok{FALSE}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{cols\_label}\NormalTok{(}\AttributeTok{total\_prop\_mean =} \StringTok{"Mean (SD)"}\NormalTok{,}
               \AttributeTok{l1\_prop\_mean =} \StringTok{"Mean (SD)"}\NormalTok{,}
               \AttributeTok{l2\_prop\_mean =} \StringTok{"Mean (SD)"}\NormalTok{,}
               \AttributeTok{concept\_prop\_mean =} \StringTok{"Mean (SD)"}\NormalTok{,}
               \AttributeTok{te\_prop\_mean =} \StringTok{"Mean (SD)"}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_stub\_indent}\NormalTok{(}\AttributeTok{rows =} \FunctionTok{everything}\NormalTok{(),}
                    \AttributeTok{indent =} \DecValTok{2}\NormalTok{) }\SpecialCharTok{|\textgreater{}} 
    \FunctionTok{tab\_style}\NormalTok{(}\FunctionTok{cell\_text}\NormalTok{(}\AttributeTok{align =} \StringTok{"left"}\NormalTok{),}
              \FunctionTok{cells\_stub}\NormalTok{())}
\end{Highlighting}
\end{Shaded}

\hypertarget{tbl-vocabulary}{}
\begin{longtable}{l|rrrrr}
\caption{\label{tbl-vocabulary}Vocabulary sizes. Total: proportion of words in both languages marked as
\emph{Understands}. L1: proportion of words in the dominant language
marked as \emph{Understands}. L2: proportion of words in the
non-cominant language marked as \emph{Understands}. Conceptual:
proportion of translation equivalents for which at least one of the
words has ben marked as \emph{Understands}. TE: proportion of
translation equivalents for which \emph{both} words have been marked as
\emph{Understands}. }\tabularnewline

\toprule
\multicolumn{1}{l}{} & \multicolumn{5}{c}{Vocabulary size (\%)} \\ 
\cmidrule(lr){2-6}
\multicolumn{1}{l}{} & Total  & L1 & L2 & Conceptual & TE \\ 
\cmidrule(lr){2-2} \cmidrule(lr){3-3} \cmidrule(lr){4-4} \cmidrule(lr){5-5} \cmidrule(lr){6-6}
\multicolumn{1}{l}{} & Mean (SD) & Mean (SD) & Mean (SD) & Mean (SD) & Mean (SD) \\ 
\midrule
\multicolumn{6}{l}{21 months} \\ 
\midrule
\hspace*{10px} Monolingual & $0.39$ ± $0.18$ & $0.53$ ± $0.18$ & $0.26$ ± $0.22$ & $0.61$ ± $0.18$ & $0.19$ ± $0.20$ \\ 
\hspace*{10px} Bilingual & $0.42$ ± $0.23$ & $0.50$ ± $0.22$ & $0.42$ ± $0.24$ & $0.56$ ± $0.21$ & $0.33$ ± $0.23$ \\ 
\midrule
\multicolumn{6}{l}{25 months} \\ 
\midrule
\hspace*{10px} Monolingual & $0.55$ ± $0.22$ & $0.76$ ± $0.14$ & $0.48$ ± $0.29$ & $0.77$ ± $0.14$ & $0.39$ ± $0.27$ \\ 
\hspace*{10px} Bilingual & $0.64$ ± $0.20$ & $0.71$ ± $0.17$ & $0.57$ ± $0.25$ & $0.75$ ± $0.20$ & $0.51$ ± $0.24$ \\ 
\midrule
\multicolumn{6}{l}{30 months} \\ 
\midrule
\hspace*{10px} Monolingual & $0.67$ ± $0.20$ & $0.81$ ± $0.16$ & $0.53$ ± $0.28$ & $0.83$ ± $0.15$ & $0.43$ ± $0.29$ \\ 
\hspace*{10px} Bilingual & $0.76$ ± $0.18$ & $0.79$ ± $0.14$ & $0.71$ ± $0.23$ & $0.85$ ± $0.10$ & $0.64$ ± $0.21$ \\ 
\bottomrule
\end{longtable}

\hypertarget{stimuli}{%
\subsection{Stimuli}\label{stimuli}}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{n\_words }\OtherTok{\textless{}{-}} \FunctionTok{length}\NormalTok{(}\FunctionTok{unique}\NormalTok{(}\FunctionTok{with}\NormalTok{(stimuli, }\FunctionTok{unlist}\NormalTok{(prime, target, distractor))))}
\end{Highlighting}
\end{Shaded}

We used 62 distinct words included in the BVQ to create the stimuli
lists. We created six stimuli lists: three in Catalan, and three in
Spanish. Each list contained 32 trials, each involving a
prime-target-distractor group. Each word played a role as either prime,
\emph{or} as target and distractor across the three lists in their
corresponding language. For instance, the Catalan word \emph{cadira}
appeared as \emph{prime} in the three lists, but never as \emph{target}
or \emph{distractor}; the Catalan word \emph{bici} appeared as
\emph{target} and \emph{distractor} across the three lists, but never as
a prime. Target-distractor pairings were held constant across the three
lists in each language. For instance, in all Catalan lists the word
\emph{bici} was paired with the word \emph{porta}. Target-distractor
pairings were also yoked, so that each member of the same
target-distractor pair appeared once as target and once as distractor in
each list. For instance, the \emph{bici}-\emph{porta} paired appeared
twice in each of the three Catalan lists: once with \emph{bici} as
target and \emph{porta} as distractor, and once with \emph{porta} as
target and \emph{bici} as distractor. This counterbalancing avoided
participants encountering looking at the target word guided solely by
that word having being named in a previous trial. Finally, prime words
appeared only once in each list: each target-distractor pair was
associated with a different prime word in both appearances. In each
list, the same prime word was presented alongside a different
target-distractor pair. For instance, the Catalan prime word
\emph{barret} was presented with the \emph{bici}-\emph{porta}
target-distractor pair in one list, with the \emph{bici}-\emph{porta}
pair in another list, and with \emph{berenar}-\emph{amanida} in the
remaining list. The order of the trials was randomised across
experimental session, so that each time a participant was tested, the
order in which the prime-target-distractor was presented was randomised.
Each participant was randomly assigned to one of the three lists in the
corresponding language (their dominant language, see
Section~\ref{sec-lp}), and always the same list across their
experimental sessions in the case of a recurrent participant.

In 16 of the 32 trials of the same list (henceforth \emph{related}
trials), the prime and the target words were phonologically related,
sharing phonological onset (at least first phoneme). In the other 16
trials (\emph{unrelated} trials), prime and target did not share
phonological onset. 8 of the 16 \emph{related} trials included a cognate
prime (\emph{cognate} trials), and the other 8 included a non-cognate
trials (\emph{non-cognate} trials). A prime word was considered cognate
if its Catalan and Spanish translation shared phonological onset.
Especial attention was paid to avoiding semantic or taxonomic
relationships between prime and target words, and between prime and
distractor words. Target and distractor word pairs were phonologically
unrelated (did not share phonological onset). Some of them shared
semantic features or a taxonomic relationship. This is the case of words
associated with especially salient referents such as animals or food. To
avoid infants guiding their gaze to these objects based on their
saliency, we paired animals and food items together. The position of the
target and distractor pictures (right or left) for each
target-distractor pair was alternated, so that in one list the target
would appear on the left, in another list it would appear on the right,
and so on.

@@

\hypertarget{auditory-stimuli}{%
\subsubsection{Auditory stimuli}\label{auditory-stimuli}}

The auditory stimuli were natural exemplars of the selected target
words, spoken by Catalan-Spanish proficient bilingual female speaker who
was instructed to pronounce each word in a toddler-directed manner.
Recordings were made with an Audio-Tecnica 328 microphone (AT2050) at a
sampling rate of 44100 Hz, in a soundproof room at the \emph{Laboratori
de Recerca en Infancia} at University Pompeu Fabra. We used the Audacity
(\protect\hyperlink{ref-reference}{\textbf{reference?}}) and Praat
(\protect\hyperlink{ref-boersma2001speak}{Boersma \& Van Heuven, 2001})
to record and edit the audio files. The speaker was presented with a
list of words in Catalan. The order of the words was pseudo-randomised,
and each word was produced three times in a row before moving to the
next word in the list. After going through all the words in the list,
the speaker went through the word list again generating three tokens for
each word, now in an inverse order (from bottom of the list to the top).
We then repeated the same procedure for the list of Spanish words. The
resulting audios were manually chunked into individual word-forms. For
each of the six tokens produced for each word, the most adequate was
selected for further processing. The audios were then transformed to
stereo by duplicating them into two channels, denoised, and finally
normalised. The mean duration of the final audios was 1.23 (\emph{SD} =
0.17) and 1.08 (\emph{SD} = 0.14) seconds for the Catalan and Spanish
lists.

@@

\hypertarget{visual-stimuli}{%
\subsubsection{Visual stimuli}\label{visual-stimuli}}

For each word, we created a picture with a typical referent. To avoid
competition between target and distractor pictures, semantically related
target-distractor pairs were perceptually distinct
(\protect\hyperlink{ref-floccia2020translation}{\textbf{floccia2020translation?}};
\protect\hyperlink{ref-arias-trejo2010}{\textbf{arias-trejo2010?}}).

@@

We defined our time window of interest from 300 ms after the onset of
the test phase (target and distractor presentation) until the end of the
test phase (2,000 ms). For each trial, we chunked the time domain into
17 time bins of 100 ms of duration. We then calculated, for each
experimental session, time bin, and condition, participant's proportion
of target and distractor fixations. Finally, we computed the empirical
logit of target fixations, which we introduced in the statistical
analyses as our response variable. Missing eye-tracker samples were
interpolated using the last-observation-carried-forward (see
\protect\hyperlink{ref-zettersten2022peekbank}{Zettersten et al., 2022}
for a similar approach).
